BERT Model Compression and Weight Sharing

### 2019-05-11 paper/paper.tex
Resolving merge conflicts, SVD

\documentclass{article}

@@

We are interested in compressing the language representation model BERT \cite{devlin2018bert}. It is the state-of-art result in many natural language processing tasks such as natural language inference, paraphrasing and named entity recognition.

\cite{cheng2017survey} provides a comprehensive overview of the state-of-art methods to compress DNNs. Of the methods summarized by the paper, the ones that are of interest to our work are weight-pruning and quantization \cite{han2015deep}, and low-rank factorization \cite{denton2014exploiting}. Knowledge distillation, first proposed by \cite{bucilua2006model} as ``knowledge transfer'', was recently popularized by \cite{hinton2015distilling}. It differs from the previously mentioned methods since it compresses an existing model by replacing it with a simpler model that is trained on its outputs.

DeepTwist, a end-to-end training framework, allows us to incorporate compression procedures into training \cite{lee2018deeptwist}. Under this framework, the model parameters are occasionally distorted during training time instead of only at the end. This framework has the benefit of simplified training procedure and faster convergence and reduced requirement for hyperparameter tuning. Experiment on MNIST dataset with DeepTwist shows that LeNet-5 and LeNet-300-100 models are able to retain predictive accuracy under high compression rate.

@@

\subsection{DeepTwist-based Weight Distortions}

Given training data $(\bm x_i, y_i)_{i=1}^N$, where $\bm x_i$ are one-hot
encoded tokens, we train by solving the standard maximum likelihood problem
where \[
\min_{\btheta} \sum_{i=1}^N L\pr{y_i, F_\btheta(\bm x_i)}
\]
for some loss function specific to the task. For instance, $L$ is cross-entropy
loss for a (multiclass) classification problem, and $L$ is mean-squared-error
for a regression problem, which correspond to log-likelihoods in Categorical and
Gaussian models, respectively. The innovation of DeepTwist-based training is
perturbing the usual gradient-based optimization by repeatedly distorting the
current weights to achieve compression objectives.

Specifically, the DeepTwist framework presented in \Cref{sec:training} applies a
distortion
step for fixed intervals during training, where \[
\btheta \gets \distortion(\btheta). 
\]
In this section, we discuss a few methods for parameterizing $\distortion$.
Generally, we may imagine a few special cases of distortion, we call a
particular distortion \emph{distortion by layer} if it is applied per layer,
i.e. \begin{align*}
\distortion'(\btheta) = (&\distortion(\bphi_1),\ldots,\distortion(\bphi_n),\\
&\distortion(\bpsi_1),\ldots, \distortion(\bpsi_m)).
\end{align*}
We call a particular distortion \emph{pretrained-only} if \[
\distortion'(\btheta) = \distortion(\bphi).
\]

@@

\paragraph{Rank Reduction} If $\bvartheta$ is a matrix, then we may apply rank
reduction via singular value decomposition \[
\bvartheta = U\Sigma V^\top
\]
where $\Sigma$ is a diagonal matrix, where $U \in $

We distort via \[
\rankreduce(\bvartheta) = \tilde U \tilde \Sigma \tilde V^\top,
\]
where $\tilde U, \tilde \Sigma, \tilde V$ are conforming matrices such that the
smallest values on the spectrum $\Sigma$ is zeroed out, as well as corresponding
values in $U,V$. If we keep only the $r$ largest spectral values, then we store
$\tilde U, \tilde \Sigma$, an $n\times r$ matrix

Example Structure:

\begin{itemize}
\item What is the formal definition of your problem?
\item What is the precise mathematical model you are using to represent it? In almost all cases this will use the probabilistic language from class, e.g.
  \begin{equation}
  z \sim {\cal N}(0, \sigma^2)\label{eq:1}
\end{equation}
But it may also be a neural network, or a non-probabilistic loss,
\[ h_t \gets \mathrm{RNN}(x_{t}, h_{t-1} )\]

This is also a good place to reference a diagram such as Figure~\ref{fig:diagram}.

\item What are the parameters or latent variables of this model that you plan on estimating or inferring? Be explicit. How many are there? Which are you assuming are given? How do these relate to the original problem description?
\end{itemize}

@@

The benefits of training in this fashion have been empirically documented
% TODO: paper on KD doing well?
and have theoretical motivation. Our goal of any network is to learn the
conditional distribution 
\[ p(y_i = k \mid \bm{x}_i).\]
In traditional training, we do this by observing the draws $y_i$ from this
distribution. However, to the extent that the teacher network prediction is a
good approximation of the true conditional distribution, the student network can
learn faster because it is being pushed toward predicting the de-noised
expectation.

>>>>

\documentclass[11pt]{article}

@@

We are interested in compressing the language representation model BERT \citep{devlin2018bert}. 
It is the state-of-art result in many natural language processing tasks 
such as natural language inference, paraphrasing and named entity recognition.

\citet{cheng2017survey} provides a comprehensive overview of the state-of-art methods to compress DNNs. Of the methods summarized by the paper, the ones that are of interest to our work are weight-pruning and quantization \cite{han2015deep}, and low-rank factorization \cite{denton2014exploiting}. Knowledge distillation, first proposed by \cite{bucilua2006model} as ``knowledge transfer'', was recently popularized by \cite{hinton2015distilling}. It differs from the previously mentioned methods since it compresses an existing model by replacing it with a simpler model that is trained on its outputs.

DeepTwist, a end-to-end training framework, allows us to incorporate compression procedures into training \citep{lee2018deeptwist}. Under this framework, the model parameters are occasionally distorted during training time instead of only at the end. This framework has the benefit of simplified training procedure and faster convergence and reduced requirement for hyperparameter tuning. Experiment on MNIST dataset with DeepTwist shows that LeNet-5 and LeNet-300-100 models are able to retain predictive accuracy under high compression rate.

@@

% \subsection{DeepTwist-based Weight Distortions}

Given training data $(\bm x_i, y_i)_{i=1}^N$, where $\bm x_i$ are one-hot
encoded tokens, we train by solving the standard maximum likelihood problem
where \[
\min_{\btheta} \sum_{i=1}^N \ell\pr{y_i, F_\btheta(\bm x_i)}
\]
for some loss function specific to the task. For instance, $\ell$ is cross-entropy
loss for a (multiclass) classification problem, and $\ell$ is mean-squared-error
for a regression problem, which correspond to log-likelihoods in Categorical and
Gaussian models, respectively. The innovation of DeepTwist-based training is
perturbing the usual gradient-based optimization by repeatedly distorting the
current weights to achieve compression objectives.

Specifically, the DeepTwist framework presented in \Cref{sec:training} applies a
distortion
step for fixed intervals during training, where \[
\btheta \gets \distortion(\btheta). 
\]
In this section, we discuss a few methods for parameterizing $\distortion$.
Generally, we may imagine a few special cases of distortion, we call a
particular distortion \emph{distortion by layer} if it is applied per layer,
i.e. \begin{align*}
\distortion'(\btheta) &= \\ (&\distortion(\bphi_1),\ldots,\distortion(\bphi_n),\\
&\distortion(\bpsi_1),\ldots, \distortion(\bpsi_m)).
\end{align*}
We call a particular distortion \emph{pretrained-only} if \[
\distortion'(\btheta) = \distortion(\bphi).
\]

@@

\paragraphdot{Rank Reduction} If $\bvartheta$ is a matrix, then we may apply
rank
reduction via singular value decomposition \[
\bvartheta = U\Sigma V^\top
\]
where $\Sigma$ is a diagonal matrix, where $U \in \R^{n\times n}, \Sigma\in \R^
{n\times n}, V\in \R^{m\times n}$, and $\bvartheta \in \R^{n\times m}$.

We distort via \[
\rankreduce(\bvartheta) = \tilde U \tilde \Sigma \tilde V^\top,
\]
where $\tilde U, \tilde \Sigma, \tilde V$ are conforming matrices such that the
smallest values on the spectrum $\Sigma$ is zeroed out, as well as corresponding
values in $U,V$. If we keep only the $r$ largest spectral values, then we store
$\tilde U \tilde \Sigma$, an $n\times r$ matrix, and $\tilde V$, an $m\times r$
matrix. Thus, instead of storing $2nm$ floats, we may store $r(n+m)$, achieving
a compression rate of \[
\rho = \frac{r(n+m)}{2mn}.
\]
Note however, that storing matrices in this manner requires an additional
matrix multiplication in evaluation time, since we have to compute the product
$\tilde U \tilde \Sigma \tilde V$.

% \subsection{Knowledge Distillation}

% Example Structure:

% \begin{itemize}
% \item What is the formal definition of your problem?
% \item What is the precise mathematical model you are using to represent it? In almost all cases this will use the probabilistic language from class, e.g.
%   \begin{equation}
%   z \sim {\cal N}(0, \sigma^2)\label{eq:1}
% \end{equation}
% But it may also be a neural network, or a non-probabilistic loss,
% \[ h_t \gets \mathrm{RNN}(x_{t}, h_{t-1} )\]

% This is also a good place to reference a diagram such as Figure~\ref{fig:diagram}.

% \item What are the parameters or latent variables of this model that you plan on estimating or inferring? Be explicit. How many are there? Which are you assuming are given? How do these relate to the original problem description?
% \end{itemize}

@@

The benefits of training in this fashion have been empirically documented
% TODO: paper on KD doing well? 
\citep{hinton2015distilling}
and have theoretical motivation. Our goal of any network is to learn the
conditional distribution 
\[ p(y_i = k \mid \bm{x}_i).\]
In traditional training, we do this by observing the draws $y_i$ from this
distribution. However, to the extent that the teacher network prediction is a
good approximation of the true conditional distribution, the student network can
learn faster because it is being pushed toward predicting the de-noised
expectation.
